### 2019-12-06 Presentation/Presentation_trial.Rmd
Rename Presentation_trial.Rmd to Presentation/Presentation_trial.Rmd

---
title: "Pre- and Post-Debate Democratic Primary Data: Twitter, Google Trends, and Polls"
subtitle: "SURV 727"
author: "Zoe Padgett and Fatou Thiam"
date: "12/6/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```